### 2017-02-26 referee-response-scientific-v1.txt
Complete referee response

I have carefully reviewed the manuscript by Casey et al. "The RAVE-on
catalog of stellar atmospheric chemical abundances..." and have the
following recommendations.

This paper presents the results of a valuable and comprehensive effort
to obtain greatly improved estimates of stellar parameters and elemental
abundances for the over 500K stars from RAVE DR5. It is clearly worthy
of publication in the ApJ, and I urge rapid acceptance.

The paper is extremely well written, and describes the procedures and
cautions in interpretation of the results very well indeed. In fact, I
have only one primary recommendation for the authors to consider.

That is, the figure I was hoping to see is one that provides the reader
a quick way to assess the distributions of the derived params and
abundances (including [Fe/H]) in the form of panels of histograms.
Basically, what I found myself wanting to see was a single set of panels
which show the full set of derived Teff, logg, [Fe/H], and [X/Fe] for
the > 500K stars that the authors have made available electronically.
This would be quite useful for other workers considering these data, and
are of course trivial to produce.

My apologies for the slight delay in responding to this manuscript, but
I was on travel, and only now carved out the time to take care of this
review. This is an extremely important paper, and one that sets a high
standard for those to follow, based on the ever-expanding data now
coming available from other surveys.

-- Author's response:

    We kindly thank the referee for their positive review of our manuscript.

    We have included the figure requested by the referee, and added a paragraph
    to the text that references this figure in the appropriate section of the
    manuscript:

        The weighted labels for all stars are shown as normalized historgrams in
        Figure \ref{fig:label-histograms}. The normalization in each axis is arbitrary, 
        since main-sequence stars do not have detailed abundances here, there are fewer
        stars with -- for example, [Ni/H] labels -- than there are stars with $\teff$
        labels. The [X/H] abundance distributions peak near Solar values, and the peaks
        in the $\teff$ and $\logg$ histograms are consistent with our expectations from
        astrophysics: the build of clump stars is visible for cool giant stars 
        ($\logg \approx 2.5$ and $\teff \approx 4750$~K), and the increase in turn-off stars
        at $\teff \approx 6000$~K is expected given the longer timescales for the turnoff.

    Unfortunately the earliest that we could reasonably include this Figure is 
    in Section 3.6 (making this new figure, Figure 6) without disturbing the
    explanatory flow of the paper.

    As a side-note: there is no need to apologise for the delay, as the lead
    author was moving internationally during this period and would have been
    unable to immediately attend to this referee response anyway.